Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-05-17 article-models-ttinternal.tex
reorganization

It would be interesting to write out a weak version of pre-universes, internal
to Martin-L\"of type theory with the function extensionality principle. 
To do this, the empty context needs to
be replaced by a contractible type, extension by dependent pair types,
judgmental equalities of terms by identifications and judgmental equalities
of families by equivalences of types. We conjecture that it is possible to
carry this out (in particular to make sure that all the constructions 
type-check). Moreover, one could then extend the notion of `weak' pre-universes
with the requirement that every internal morphism is weakly anodyne. This
should give an internal theory of weak higher groupoids.

@@

>>>>

It would be interesting to write out a weak version of pre-universes, internal
to Martin-L\"of type theory with the function extensionality principle. 
To do this, the empty context needs to
be replaced by a contractible type, extension by dependent pair types,
judgmental equalities of terms by identifications and judgmental equalities
of families by equivalences of types. We conjecture that it is possible to
carry this out (in particular to make sure that all the constructions 
type-check). This could serve as a starting point for investigating internal
models without truncatedness assumptions and for investigating internal higher
categories. 
Moreover, one could then extend the notion of `weak' pre-universes
with the requirement that every internal morphism is weakly anodyne. This
could give an internal theory of weak higher groupoids.

@@

\begin{thm}
Let $(A,P,\epsilon_0,\epsilon_1,\phi_0,\phi_1)$ be an extension-empty algebra
in context $\Gamma$. Then the sextuple
\begin{equation*}
(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1},\phi_1,
\jcomp{}{\epsilon_0}{\phi_1})
\end{equation*}
is an extension-empty algebra in context $\ctxext{\Gamma}{A}$. 
\end{thm}

\begin{proof}
We first check whether the quadruple $(P,\jcomp{}{\epsilon_0}{P},\phi_1,\jcomp{}{\epsilon_0}{\phi_1})$
is an empty algebra in context $\ctxext{\Gamma}{A}$. Thus we need to verify
that $\subst{\phi_1}{\jcomp{}{\epsilon_0}{P}}\jdeq P$ and that
$\subst{\phi_1}{\jcomp{}{\epsilon_0}{\phi_1}}\jdeq\phi_1$. Both these judgmental
equalities follow from \autoref{lem:empalg-mor}.

To complete the proof, we need to verify that
\autoref{extempalg-eq1,extempalg-eq2,extempalg-eq3,extempalg-eq4} hold in our
current setting.
\end{proof}
